---
title: "Background"
description: |
  Problem description and pre-existing methodologies.
bibliography: biblio.bib  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In Mexico, the elections take place on a Sunday, and the official results 
are presented to the population several days after the election day. To prevent
unjustified victory claims during that period, the electoral authority organizes 
a quick count the same day of the election. With highly competed electoral 
processes, the quick count has become very important, the rapidity and precision 
of its results auspicious an environment of trust, and it serves as a tool 
against fraud.

The quick count consists in selecting a random sample of the polling stations 
and estimating the percentage of votes in favor of each candidate. The electoral
authority creates a committee of specialists in charge of the quick count, whose
responsibilities encompass, mostly, the sample design, and the operation of 
statistical methods to produce the inferences. The inferences are presented by
probability intervals with an associated probability of at least 0.95.

The information system starts at 6 p.m. and, every 5 minutes, collects all the 
sample information sent. Thus, the system produces a sequence of accumulative 
files used to determine the available percentage of the sample and its 
distribution over the country. The partial samples are analysed with the 
estimation methods to track the trend of the results. The quick-count results 
are made public using one of these partial samples since the complete sample 
takes too long to arrive. The committee reports a result when certain conditions 
are met, such as the arrival of a large part of the sample and the stability in 
the estimates of the last partial samples.

The context of the quick count estimation problem is then as follows. We have a
stratified sample designed by the committee, so we know the inclusion 
probabilities and the strata weights. The key challenge is that we have to 
estimate with incomplete samples, which may imply limited (or null) sample size 
in some strata and where the missingness is not at random. In fact, the 
probability that a polling station is not reported at a certain time depends 
heavily on the response we intend to measure.


## Existing approaches

We will use as example the 2018 national counts, where the strata were defined
by the federal districts (an electoral spatial unit, defined with states).

```{r}
library(tidyverse)
library(quickcountmx)

conteo_2018 <- conteo_2018 %>%
    mutate(edo_distrito = paste(ID_ESTADO, ID_DISTRITO, sep = "-"))
stratum_sizes <- conteo_2018%>% 
    group_by(edo_distrito) %>%
    summarise(n_stratum = n())

glimpse(stratum_sizes)
```

Sample allocation was proportional to strata size.

```{r}
sample_2018 <- select_sample_prop(conteo_2018, stratum = edo_distrito, 0.06)
```

### Ratio estimator

The combined ratio estimator is a frequently used approach. For a given candidate
the estimate $\hat{p}$ is:

$$\hat{p}=\frac{\sum_h \frac{N_h}{n_h} \sum_i Y_{hi}}{\sum_h \frac{N_h}{n_h} \sum_i X_{hi}}$$
where:

* $Y_{hi}$ total number of votes for the candidate in the $i$th polling station, 
belonging to the $h$th stratum.

* $X_{hi}$ total number of votes in the $i$th polling station, 
belonging to the $h$th stratum. 

* $N_h$ total number of polling stations in the $h$th stratum.

* $n_h$ sample size for the $h$th stratum.

With ths selected sample size we compute estimates using ratio estimator, and
standard errors using bootstrap.

```{r}
ratio_2018 <- ratio_estimation(sample_2018, stratum = edo_distrito,
  data_stratum = stratum_sizes, n_stratum = n_stratum, 
  any_of(c("AMLO", "JAMK", "RAC", "CAND_IND_01", "CAND_IND_02")))

ratio_2018 %>% 
  mutate(across(where(is.numeric), round, 3))
```

### Bayesian Normal Model

In [@mendoza2016] the authors propose to fit a model independently for each 
candidate in each stratum, in the national election this would account for 
1,500 models. For a given candidate, on a given stratum, the model is defined as
follows:

* **Likelihood**: let $X_{i}$ be the votes for the  candidate in the 
$i$th polling station, then

$$X_{i}\big|\theta,\tau\sim N\bigg(n_i\theta, \frac{\tau}{n_i}\bigg)$$

For $i = 1,...,c$, where $c$ is the number of polling stations in the stratum,
$n_i$ stands for the size of the nominal list (list of registered voters) in 
the poll, $\tau$ is the 
precision. Note that for each candidate, $\tau$ is assumed to be constant within 
the corresponding stratum (and unrelated to $\theta$), and vote counts are 
also assumed to be independent across candidates of the same stratum.

* Prior distributions

$$p(\theta,\tau)\propto \tau^{-1}I(\tau>0)I(0<\theta<1)$$

* Posterior distribution

$$p(\theta, \tau|X) \sim N\bigg(\theta \bigg| \frac{\sum_{i=1}^{c}x_{i}}{\sum_{i=1}^{c}n_{i}}, \tau\sum_{i=1}^{c}n_i\bigg)I(0<\theta<1)\times Ga\bigg(\tau\bigg|\frac{c-1}{2}, \frac{1}{2}\bigg[\sum_{i=1}^{c}\frac{x_{i}^2}{n_i}-\frac{\big(\sum_{i=1}^{c}x_{i}\big)^2}{\sum_{k=1}^{c_i}n_i}\bigg]\bigg)$$
Finally, let us notice that $\theta$ (let us include the candidate for this, 
$\theta_j$ for the 
$j$th candidate) is defined as the proportion of people in the nominal list of
the stratum whose preference is for the $\j$th candidate. However the parameters 
of interest are not the $\theta_j$, since we
are looking for the effective proportions of votes, so the authors define the 
parameters of interest as:

$$\lambda_{j}= \frac{\theta_j}{\sum_k \theta_k}$$

We implement the model, and run for the same sample as ratio estimator:

```{r}
sample_2018 <- sample_2018 %>% 
  mutate(LISTA_NOMINAL_CASILLA = ifelse(LISTA_NOMINAL_CASILLA == 0, 750,
                                        LISTA_NOMINAL_CASILLA))
nnp_2018 <- nnp_estimation(data = sample_2018, ln = LISTA_NOMINAL_CASILLA,
  stratum = edo_distrito, data_stratum = stratum_sizes, n_stratum = n_stratum, 
  any_of(c("AMLO", "JAMK", "RAC", "CAND_IND_01", "CAND_IND_02")), n_sims = 100)
nnp_2018$lambdas_summary
```

